Synthesizing Symmetric Lenses

### 2019-02-26 papers/SymmetricLenses/main.tex
edits to abstract

\hspace*{1em}We show how to extend this synthesis algorithm to a much wider class of
lenses, which we call {\em simple symmetric lenses}, including not only all
of the bijective lenses and the more widely used ``asymmetric''
lenses, but also a rich subset of the full-blown
``symmetric lenses'' proposed by Hofmann \ETAL.

Synthesizing simple symmetric lenses is substantially more challenging than
synthesizing bijective lenses: Since some of the information on each side
can be ``disconnected'' from the other side, there will, in general, be {\em
  many} lenses that agree with a given example. To guide the search process,
we use {\em stochastic regular expressions} and ideas from information
theory to estimate the amount of information propagated by a candidate lens,
in conjunction with user annotations identifying parts of the source and
target data structures as either {\em irrelevant} or {\em essential}.

Simple symmetric lenses are also of independent theoretical interest, being
the largest class of symmetric lenses that do not rely on persistent
internal state.

>>>>

\hspace*{1em}We show how to extend this synthesis algorithm to a much wider class of
lenses, which we call {\em simple symmetric lenses}, including not only all
of the bijective lenses and the more widely used ``asymmetric''
lenses, but also a rich subset of the full-blown
``symmetric lenses'' proposed by Hofmann \ETAL  
Intuitively, simple symmetric lenses allow some information to be present
on one side but not the other and vice versa.
Simple symmetric lenses are of independent theoretical interest, being
the largest class of symmetric lenses that do not rely on persistent
internal state.  
Synthesizing simple symmetric lenses is substantially more challenging than
synthesizing bijective lenses: Since some of the information on each side
can be ``disconnected'' from the other side, there will, in general, be {\em
  many} lenses that agree with a given example. To guide the search process,
we use {\em stochastic regular expressions} and ideas from information
theory to estimate the amount of information propagated by a candidate
lens, generally prefering lenses that propagate more information, 
as well as user annotations identifying parts of the source and
target data structures as either {\em irrelevant} or {\em essential}.
